---
title: 'Analyzing Centre for Disease Control(CDC) Datasets'
author: "Sujit Krishnankutty"
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}

load("C:/Users/sujit/Downloads/cdc1.Rdata")

# View(cdc1)
cdc1$exerany = factor(cdc1$exerany)

#Categorizing people based on the exercise

levels(cdc1$exerany)=c("No_Exercise", "Exercise")

#Categorizing people based on the smoker/non-smoker

cdc1$smoke100 = factor(cdc1$smoke100)

levels(cdc1$smoke100) =c("Smoke_less", "Smoke_more")

cdc1$hlthplan = factor(cdc1$hlthplan)
levels(cdc1$hlthplan) =c("No_Health_cover", "Health_cover")


```

#Tabular format for finding male-female proportion for excellent and very-good health category

```{r}
prop.table(table(cdc1$genhlth=="excellent"| cdc1$genhlth=="very good",cdc1$gender),2)

prop.table(table(cdc1$genhlth=="excellent"| cdc1$genhlth=="very good",cdc1$gender),2)[4]


```
From these proportions, we can conclude that Men have better health than women.

```{r}

smokers<-subset(cdc1,cdc1$smoke100=='Smoke_more')

prop.table(table(smokers$genhlth=="excellent"| smokers$genhlth=="very good",smokers$gender),2)

prop.table(table(smokers$genhlth=="excellent"| smokers$genhlth=="very good",smokers$gender),2)[2]
```
```{r}
#Health-wise segregation for Smoking category

prop.table(table(smokers$genhlth=="excellent"| smokers$genhlth=="very good",smokers$gender),2)[4]
```
From the above two values we can say that in the category where people smoke more, men have better health than women i.e. smoker's category.


```{r}
#Health-wise segregation for non-Smoking category
non_smokers<-subset(cdc1,cdc1$smoke100=='Smoke_less')
prop.table(table(non_smokers$genhlth=="excellent"| non_smokers$genhlth=="very good",non_smokers$gender),2)

prop.table(table(non_smokers$genhlth=="excellent"| non_smokers$genhlth=="very good",non_smokers$gender),2)[2]
```
For the non-smokers category too, we observe men have better health than women.




Based on the above results we can say that men have better health compare to women
Health of non-smokers( 0.6964286 = m, 0.6 = f ) seems to be better than smokers with ( 0.6 = m, 0.5454545 = f).

we can conclude that men have better health than women in both the cases.



```{r}
skew <-function(x)
{
  if(is.numeric(x))
  {
    n <- length(x)
    
    m3<-sum(((x-mean(x))^3)/n)
    m2<-sum(((x-mean(x))^2)/n)
    (m3/(m2^(3/2)))
  }
  else
  {
    NA
  }  
}
```


```{r}
skew(mtcars$wt)
```


d.
```{r}
rep <- replicate(1000,skew(rnorm(50,mean=0,sd=1)))

hist(rep)
```
```{r}
mean(rep)

```
e.
```{r}

skew(rexp(50,rate=1))

```

```{r}
rep2 <- replicate(1000,skew(rexp(50,rate=1)))
hist(rep2)
```

```{r}
mean(rep2)
```


```{r}
boxplot(rep,rep2)
```

It can be observed that the mean of exponential distribution is higher as compared to that of the rnorm (random normal distribution).
Histogram of normal distribution is mean-centred at zero & for exponential histogram is skewed left.
There are higher outliers in case of exponential, whereas in rnorm we have both lower as well as higher outliers.

Finding skewness for every variable of CDC dataset
```{r}

skewness<-sapply(cdc1,skew)
skewness

```

Plotting weight-weight_desired vs age

```{r}
plot(cdc1$weight-cdc1$wtdesire~cdc1$age ,pch=20 ,col=
  ifelse(cdc1$weight>cdc1$wtdesire & cdc1$age>30,"red",
    ifelse(cdc1$weight>cdc1$wtdesire & cdc1$age<=30,"blue",
      ifelse(cdc1$weight<=cdc1$wtdesire & cdc1$age>30,"orange",
        ifelse(cdc1$weight<=cdc1$wtdesire & cdc1$age<=30,"cyan","black")))), xlab = "Age", ylab = "Weight- Weight_desire")
```

```{r}
boxplot.out<-function(x)
{
  n<-length(boxplot.stats(x)$out)
  LQ<-fivenum(x)[2]
  HQ<-fivenum(x)[4]
  IQR<-HQ-LQ
  high<-0
  low<-0
  for(i in 1:n)
  {
    if(n!=0)
    {
      if(boxplot.stats(x)$out[i]<(LQ-(1.5*IQR)))
      {
        low<-low+1
        
      }
      else if(boxplot.stats(x)$out[i] > (HQ+(1.5*IQR)))
      {
        high<-high+1
      
      }
    }
  }
  ifelse(high!=0 & low!=0 ,"both",
         ifelse(high==0 & low!=0, "low",
                ifelse(high !=0 & low==0,"high","none")))
}
```

Plot for (weight-wtdesire) vs genhlth

```{r}
boxplot(weight-wtdesire ~ genhlth, data = cdc1)
```

d.
```{r}
Outlier_type<-tapply(cdc1$weight-cdc1$wtdesire,cdc1$genhlth,boxplot.out)
Outlier_type 

```

e.
```{r}
 
color_vector<-NULL

for (i in 1:length(Outlier_type))
{
  if(Outlier_type[i]=="both")
  {
    color_vector<-c(color_vector,"red")
  }
  else if(Outlier_type[i]== "high")
  {
    color_vector<-c(color_vector,"blue")
  }
  else if(Outlier_type[i]=="low")
  {
    color_vector<-c(color_vector,"green")
  }
  else
  {
    color_vector<-c(color_vector,"yellow")
  }
}
boxplot(weight-wtdesire~genhlth,data=cdc1,col=color_vector)
```




